\chapter{System and Benchmark Artifacts}

This appendix summarizes the implementation and benchmark artifacts used by the thesis. These artifacts support reproducibility of the engineering evaluation, but they do not establish clinical validity.

\section{System Implementation Notes}

The backend exposes chat and streaming routes through the API layer. The dialogue graph is implemented in the AI engine and uses a shared blackboard state. The frontend provides the chat interface, conversation history, model selection, and DASS-21 onboarding form.

The system response format separates internal drafts from visible messages. Peer drafts can be logged for debugging, but only therapist-approved messages are sent to the user. This is important for both safety and benchmark consistency because the conversation history should contain only what the user actually saw.

\section{Benchmark Artifacts}

Each VSM benchmark run writes structured files:

\begin{itemize}
    \item \texttt{per\_case\_results.jsonl}: raw outputs, metadata, and deterministic scores for each case;
    \item \texttt{failures.jsonl}: hard failures, route mismatches, stage mismatches, peer mismatches, and fallback events;
    \item \texttt{judge\_results.jsonl}: judge scores for each turn;
    \item \texttt{score\_summary.json}: aggregate benchmark metrics;
    \item \texttt{human\_audit\_template.csv}: sampled turns for manual review;
    \item \texttt{report.md}: generated benchmark tables and figures.
\end{itemize}

\section{Recommended Final Evaluation Procedure}

The final benchmark should be run in stages. First, the full Probe split should be run with the proposed system and lightweight baselines. Second, a balanced Session-Core subset should be evaluated with a real LLM judge. Third, 20--30 sampled turns should be manually audited. Only after these steps should the full Session-Core split be used for final reporting.

This staged procedure reduces the risk of spending API budget on a flawed benchmark configuration. It also makes the final report easier to defend because each result can be traced to raw outputs, deterministic checks, judge scores, and manual audit notes.
